---
title: "Combine population estimates and population projections"
author: "Joel Barnes"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document: 
    toc: true
    number_sections: false
    toc_float: true
    toc_depth: 5
  word_document: 
    toc: yes
    toc_depth: 5
  pdf_document:
    toc: true
    number_sections: true
    toc_float: true
    toc_depth: 5
editor_options: 
  chunk_output_type: console
---

<style>
  html, body { font-size: 14px; }
  h1 { font-size: 1.75rem; margin-top: 30px; }
  h1.title { font-size: 2rem; color: #484848; }
  h2 { font-size: 1.35rem; margin-top: 30px; }
  h3 { font-size: 1.15rem; margin-top: 30px; }
  h4 { font-size: 1.05rem; margin-top: 30px; }
  h4.date { margin-top: 0rem; margin-bottom: 3rem; }
  a, a:hover { color: #fff; }
  p.caption { margin-top: 2rem; }
  blockquote { border-left: 5px solid #cccccc; font-size: 1rem; background-color: #00AFBB; color: #ffffff; font-weight: bold; }
  .interpretation blockquote { border-left: 5px solid #cccccc; font-size: 1rem; background-color: #00AFBB; color: #ffffff; font-weight: normal; }
  .gmisc_table, table { width: 100%; }
  img { width: 100%; height: auto; }
  .list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
    background-color: #00AFBB;
    border-color: #00AFBB;
}
</style>

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

> Note: this document is confidential. Please do not share.

## Load dependencies

```{r}
library(dplyr); library(readr)
```

## Import data

```{r}
# CANSIM Table 17-10-0005-01 (population estimates)
d <- read_csv("c:/users/joelb/onedrive/github/cansim/CANSIM Table 17-10-0005-01 - wrangled.csv")
  
# CANSIM Table 17-10-0057-01 (population projections)
d2 <- read_csv("c:/users/joelb/onedrive/github/cansim/CANSIM Table 17-10-0057-01 - wrangled.csv")
```

## Bind imported data

```{r}
d3 <- bind_rows(d, d2)
```

## Export data

```{r}
write.csv(d3, "c:/users/joelb/onedrive/github/cansim/CANSIM population data - estimates and projections combined.csv", na = "", row.names = FALSE)
```