---
title: "Wrangle CANSIM Table 17-10-0057-01"
author: "Joel Barnes"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document: 
    toc: true
    number_sections: false
    toc_float: true
    toc_depth: 5
  word_document: 
    toc: yes
    toc_depth: 5
  pdf_document:
    toc: true
    number_sections: true
    toc_float: true
    toc_depth: 5
editor_options: 
  chunk_output_type: console
---

<style>
  html, body { font-size: 14px; }
  h1 { font-size: 1.75rem; margin-top: 30px; }
  h1.title { font-size: 2rem; color: #484848; }
  h2 { font-size: 1.35rem; margin-top: 30px; }
  h3 { font-size: 1.15rem; margin-top: 30px; }
  h4 { font-size: 1.05rem; margin-top: 30px; }
  h4.date { margin-top: 0rem; margin-bottom: 3rem; }
  a, a:hover { color: #fff; }
  p.caption { margin-top: 2rem; }
  blockquote { border-left: 5px solid #cccccc; font-size: 1rem; background-color: #00AFBB; color: #ffffff; font-weight: bold; }
  .interpretation blockquote { border-left: 5px solid #cccccc; font-size: 1rem; background-color: #00AFBB; color: #ffffff; font-weight: normal; }
  .gmisc_table, table { width: 100%; }
  img { width: 100%; height: auto; }
  .list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
    background-color: #00AFBB;
    border-color: #00AFBB;
}
</style>

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

> Note: this document is confidential. Please do not share.

## Load dependencies

```{r}
library(cansim); library(dplyr); library(stringr)
```

## Import data

```{r eval = FALSE}
# CANSIM Table 17-10-0057-01 (projected population)
d <- get_cansim("17-10-0057-01")
```

```{r echo = FALSE}
d <- readRDS("c:/users/joelb/onedrive/github/cansim/CANSIM Table 17-10-0057-01.rds")
```

## Wrangle data
### Filter projection scenario

```{r}
# Retain M3 scenario
d <- d %>% filter(str_detect(`Projection scenario`, "M3"))
```

### Filter on age groups

```{r}
# Remove single years and retain groupings
d <- d %>% filter(str_detect(tolower(`Age group`), "all|over|to"))
```

### Remove age groups under 20 years old

```{r}
d <- d %>% 
  filter(
    ! `Age group` %in% paste0(c(0, 5, 10, 15), " to ", c(4, 9, 14, 19), " years")
  )
```

### Standardize age group values

```{r}
# Coerce to character class
d$`Age group` <- as.character(d$`Age group`)

# Standardize values
d$`Age group` <- sapply(d$`Age group`, function(x) {
  if(is.na(x)) {
    output <- NA
  } else if(str_detect(tolower(x), "all")) {
    output <- "All"
  } else if(str_detect(x, "over")) {
    output <- "[100,max["
  } else {
    output <- paste0(
      "[",
      strsplit(x, " ")[[1]][1],
      ",",
      as.integer(strsplit(x, " ")[[1]][3]) + 1,
      "["
    )
  }
})
```

### Standarize sex values

```{r}
# Coerce to character class
d$Sex <- as.character(d$Sex)

# Standardize values
d$Sex <- sapply(d$Sex, function(x) {
  if(is.na(x)) {
    output <- NA
  } else if(x == "Both sexes") {
    output <- "All"
  } else {
    output <- substr(x, 1, nchar(x) - 1)
  }
})
```

### Filter on year

```{r}
# Retain years 2022 to 2043 (missing values in years beyond 2043)
d <- d %>% filter(REF_DATE >= 2022, REF_DATE <= 2043)
```

### Filter on province

```{r}
# Remove the territories
d <- d %>% filter(
  ! GEO %in% c("Northwest Territories", "Nunavut", "Yukon")
)
```

### Standardize province values

```{r}
d$GEO <- sapply(d$GEO, function(x) {
  if(is.na(x)) {
    output <- NA
  } else if(x == "Canada") {
    output <- "All"
  } else if(x == "Newfoundland and Labrador") {
    output <- "Newfoundland"
  } else if(x == "Prince Edward Island") {
    output <- "P.E.I."
  } else {
    output <- x
  }
})
```

### Filter on vectors

```{r}
# Retain select vectors
d <- d %>% select(
  REF_DATE,
  GEO,
  `Age group`,
  Sex,
  val_norm
)
```

### Rename vectors

```{r}
names(d) <- c(
  "year",
  "province",
  "age",
  "sex",
  "population_count"
)
```

### Add engine vector

```{r}
d$engine <- "CANSIM"
```

### Rearrange vectors

```{r}
d <- d[c("engine", "province", "sex", "year", "age", "population_count")]
```

### Recompute aggregate population counts

```{r}
# Add a couple of vectors
d$population_count_pohem <- NA
d$all_count <- NA

# Count, by row, the number of vectors that factor into the aggregated population count
for(i in 1:nrow(d)) {
  d$all_count[i] <- sum(unlist(d[i, c("province", "sex", "year", "age")]) == "All")
}

# Compute population counts for rows without aggregated data
for(i in which(d$all_count == 0)) {
  d$population_count_pohem[i] <- sum(unlist(d$population_count[d$province == d$province[i] & d$sex == d$sex[i] & d$year == d$year[i] & d$age == d$age[i]]))
}

# Compute aggregated population counts for rows with one-factor aggregation 
for(i in which(d$all_count == 1)) {
  if(d$province[i] == "All") {
    d$population_count_pohem[i] <- sum(unlist(d$population_count[d$province != "All" & d$sex == d$sex[i] & d$year == d$year[i] & d$age == d$age[i]]))
  } else if(d$sex[i] == "All") {
    d$population_count_pohem[i] <- sum(unlist(d$population_count[d$province == d$province[i] & d$sex != "All" & d$year == d$year[i] & d$age == d$age[i]]))
  }  else if(d$age[i] == "All") {
    d$population_count_pohem[i] <- sum(unlist(d$population_count[d$province == d$province[i] & d$sex == d$sex[i] & d$year == d$year[i] & d$age != "All"]))
  }
}

# Compute aggregated population counts for rows with two-factor aggregation 
for(i in which(d$all_count == 2)) {
  if(d$province[i] == "All" & d$sex[i] == "All") {
    d$population_count_pohem[i] <- sum(unlist(d$population_count[d$province != "All" & d$sex != "All" & d$year == d$year[i] & d$age == d$age[i]]))
  } else if(d$province[i] == "All" & d$age[i] == "All") {
    d$population_count_pohem[i] <- sum(unlist(d$population_count[d$province != "All" & d$sex == d$sex[i] & d$year == d$year[i] & d$age != d$age[i]]))
  } else if(d$sex[i] == "All" & d$age[i] == "All") {
    d$population_count_pohem[i] <- sum(unlist(d$population_count[d$province == d$province[i] & d$sex != "All" & d$year == d$year[i] & d$age != "All"]))
  } 
}

# Compute aggregated population counts for rows with three-factor aggregation 
for(i in which(d$all_count == 3)) {
  d$population_count_pohem[i] <- sum(unlist(d$population_count[d$province != "All" & d$sex != "All" & d$year == d$year[i] & d$age != "All"]))
}

# Add a vector the computes the relative difference
d$relative_difference <- (d$population_count_pohem - d$population_count) / d$population_count * 100
```

### Add rows for 0-19 year-olds

```{r}
# Create grid of unique values by province, sex and year
grid <- expand.grid(
  province = unique(d$province),
  sex = unique(d$sex),
  year = unique(d$year)
)

# Iterate through the grid
for(i in 1:nrow(grid)) {
  d <- bind_rows(
    d,
    tibble(
      engine = "CANSIM",
      province = grid$province[i],
      sex = grid$sex[i],
      year = grid$year[i],
      age = "[min,20[",
      population_count = 0,
      population_count_pohem = 0,
      all_count = NA,
      relative_difference = NA
    )
  )
}

```

## Export data

```{r}
write.csv(d, "c:/users/joelb/onedrive/github/cansim/CANSIM Table 17-10-0057-01 - wrangled.csv", na = "", row.names = FALSE)
```